---
title: "R Markdown and Plotly: Finnish Traffic Accidents 2015"
author: "vtenhunen"
date: "30 October 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
# libraries
library(plotly)
library(dplyr)
library(knitr)
library(lubridate)
library(ggplot2)

# make sure that we have right locale here
Sys.setlocale("LC_TIME","en_US.UTF-8")

# Some knitr chunk options
knitr::opts_chunk$set(echo = FALSE)
```

```{r getdata}
# Let's have the Finnish road accident data
DataURL="https://www.avoindata.fi/dataset/35f439f6-4512-444f-afd1-444356cb9524/resource/df40544f-7967-4224-8f02-6e0543f2e623/download/Tieliikenneonnettomuudet2015.zip"
DataFile="./RoadTrafficAccidentsIn2015.zip"
      
      if(!file.exists(DataFile)){
            # Get the data for the assignment
            download.file(DataURL, DataFile)
      }

# It is a zip packet, so we unzip it
      unzip(DataFile)
```

```{r readdata}

# Next we read the data to the data frame 
alldata = read.csv("./Tieliikenneonnettomuudet_2015/tieliikenneonnettomuudet_2015_onnettomuus.csv", na.strings="NA", sep = ";", header = TRUE, fileEncoding = "latin1")

# let's take only some columns:
# Onnett_id, X, Y, Kuolleet, Loukkaantuneet, 
accdata <- select(alldata, Onnett_id, Vuosi, Kk, Päivä, Tunti, Vakavuusko, X, Y, Kuolleet, Loukkaant)

# Remove NA-information
accdata <- accdata[complete.cases(accdata),]

# Now we convert the day format to the standard format 
accdata$Päivä <- as.Date(parse_date_time(accdata$Päivä,"dmy"))

### Plot 3
## Here we calculate accident frequenses, deads and injured by hour
accbyhour <- accdata %>%
  group_by(Tunti) %>%
  summarise(freq = n(), dead = sum(Kuolleet), injured = sum(Loukkaant)) 
# clean -1 hour away, don't know what it is
accbyhour<-accbyhour[!(accbyhour$Tunti=="-1"),]

### Plot 2
## Here we calculate accident frequenses, deads and injured by day
accbyday <- accdata %>%
  group_by(Päivä) %>%
  summarise(freq = n(), dead = sum(Kuolleet), injured = sum(Loukkaant)) 
# Rename columns
colnames(accbyday)[1] <- "Day"
colnames(accbyday)[2] <- "Accidents"

### Plot 1
## Here we calculate accident frequenses, deads and injured by month
accbymonth <- accdata %>%
  group_by(Kk) %>%
  summarise(freq = n(), dead = sum(Kuolleet), injured = sum(Loukkaant)) 

```


## Presentation

This presentation include following graphs about Finnish traffic accident in 2015:

- Traffic accidents, deads and injured by month
- Traffic accidents by day of the year
- Traffic accidents, deads and injured by hour of the day
- About the data

All plots has been created with the plotly. Source code is availbale in GitHub: http://

## Accidents, dead and injured by month

```{r accbymonthlayered, echo=FALSE, eval=TRUE}

#First create the plotly and after that add traces and make some texts
plot_ly(accbymonth, x = ~Kk, y = accbymonth$freq, name = 'Accidents', type = 'scatter', text = paste("Month: ", accbymonth$Kk, "/2015", "<br>Accidents: ", accbymonth$freq), mode = 'lines+markers') %>%
      add_trace(y = ~injured, name = 'Injured', text = paste("Month: ", accbymonth$Kk,"/2015", "<br>Injured: ", accbymonth$injured), mode = 'lines+markers') %>%
      add_trace(y = ~dead, name = 'Dead', text = paste("Month: ", accbymonth$Kk, "/2015", "<br>Dead: ", accbymonth$dead), mode = 'lines+markers')

```

## Accidents by day
```{r accidentfreq, echo = FALSE, eval=TRUE}

# Let's use here ggplotly
p <- ggplot(accbyday, aes(x=Day, y=Accidents)) +
    geom_point(shape=1, color="blue") + 
    geom_smooth(method=lm) + 
      labs(x="Day", y="")
# print the result
ggplotly(p)

```


## Accidents, deads and injured by hour of the day
```{r accidentsbyday, echo = FALSE, eval=TRUE}

accbyhour$Tunti <- factor(accbyhour$Tunti, levels = accbyhour[["Tunti"]])

p <- plot_ly(accbyhour, x = ~Tunti, y = ~freq, type = 'bar', name = 'Accidents by hour of the day', marker = list(color = 'rgb(49,130,189)')) %>%
        add_trace(y = ~accbyhour$injured, name = 'Injured by hour', marker = list(color = 'rgb(139,49,204)')) %>%
  add_trace(y = ~accbyhour$dead, name = 'Dead by hour', marker = list(color = 'rgb(255,0,0)')) %>%
  layout(xaxis = list(title = "Hour", tickangle = -45),
         yaxis = list(title = ""),
         margin = list(b = 100),
         barmode = 'group')
p
```

## About the data

- Finnish Transport Agency collects annual road trafic accident data
- It is based on information received from the law enforcement officials and Statistics Finland
- Source: http://tinyurl.com/zk2fshn
- Licence: CC 4.0 BY (https://creativecommons.org/licenses/by/4.0/)

